\textbf{Evaluating RISC-V Enclaves:}\\ Security Without the Slowdown?

### 2025-09-02 draft_presentation.tex
new diagram

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{fontawesome5} % for icons
\usepackage{pifont}       % for checkmarks
\usepackage{xcolor}       % for color definitions
\usepackage{colortbl}     % for \rowcolor, \rowcolors, \cellcolor in tables
\usepackage{tikz}         % for drawing graphics
\usetikzlibrary{patterns,positioning} % enable patterns and positioning for tikz
\usepackage{pgfplots} 
%\pgfplotsset{compat=1.17} % set PGFPlots compatibility

@@

% Title and Author
\title[Evaluating RISC-V Enclaves]{\textbf{Evaluating RISC-V Enclaves:}\\ Security Without the Slowdown?}
\author{\textbf{Basil Ugbomoiko}}
\institute[UzL]{Universität zu Lübeck \\ Institut für Technische Informatik}
\date{Master’s Thesis Defense — \today}
% Add logo (adjust path as needed)
\titlegraphic{\hfill\includegraphics[height=1.5cm]{uzl-logo.pdf}}
\begin{document}

@@

\begin{frame}{Problem Statement: Challenge}
\begin{itemize}
    \item[\faExclamationTriangle] Traditional security assumes a trustworthy OS — \alert{assumption broken}.
    \item[\faLock] Need isolation of critical code/data \textbf{even if the OS is compromised}.
\end{itemize}
\vfill
%\scriptsize\textit{This sets the stage for Trusted Execution Environments (TEEs).}
\end{frame}

@@

\begin{frame}{Benchmarking Tools}
    \small
    \begin{itemize}
        \item \textbf{Dhrystone:} Measures CPU performance (DMIPS). \pause
        \item \textbf{CoreMark:} Focuses on integer performance, used in embedded systems benchmarking. \pause
        \item \textbf{Two configurations:}
        \begin{itemize}
            \item \textbf{Native Execution:} Benchmark runs as a normal process. \pause
            \item \textbf{Enclave Execution:} Benchmark runs within a Keystone enclave with memory isolation. \pause
        \end{itemize}
    \end{itemize}
    \vspace{0.5em}
    \uncover<5->{\textbf{Key Focus:} These tools provide insights into both CPU-bound and system-level performance.}
\end{frame}

@@

% Legend
\begin{scope}[xshift=0cm,yshift=-1.5cm]
\draw[fill=keystoneblue!60] (0,0) rectangle (0.4,0.4) node[right=0.5cm] {Actual region (contiguous)};
\draw[fill=warmaccent!60] (5,0) rectangle (5.4,0.4) node[right=0.5cm] {Actual region (fragmented)};
\draw[pattern=north east lines,pattern color=black] (0,-0.6) rectangle (0.4,-0.2) node[right=0.5cm] {Wasted padding};
\draw[fill=gray!50,opacity=0.4] (5,-0.6) rectangle (5.4,-0.2) node[right=0.5cm] {PMP entry count bar};
\end{scope}

@@

% Thank You Slide with visual focus
\begin{frame}
\centering
\Huge\textbf{Thank You}\\
\large Questions? \faQuestionCircle
\vfill
\small Universität zu Lübeck — Master’s Thesis Defense
\end{frame}
\end{document}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{fontawesome5} % for icons
\usepackage{pifont}       % for checkmarks
\usepackage{xcolor}       % for color definitions
\usepackage{colortbl}     % for \rowcolor, \rowcolors, \cellcolor in tables
\usepackage{tikz}         % for drawing graphics
\usetikzlibrary{patterns,positioning} % enable patterns and positioning for tikz
\usepackage{pgfplots}
\usepackage[transitions={dissolve}]{pdfpc}
%\pgfplotsset{compat=1.17} % set PGFPlots compatibility

@@

% Title and Author
\title[Evaluating RISC-V Enclaves]{\textbf{Evaluating RISC-V Enclaves:}\\ Security Without the Slowdown?}
\subtitle{Performance Evaluation and Optimization of Keystone Enclaves on RISC-V}
\author{\textbf{Basil Ugbomoiko}}
\institute[UzL]{Universität zu Lübeck \\ Institut für Technische Informatik}
\date{Master’s Thesis Defense — \today}
% Add logo (adjust path as needed)
\titlegraphic{\hfill\includegraphics[height=1.5cm]{uzl-logo.pdf}}
\begin{document}

@@

\begin{frame}{Problem Statement: Challenge}
\begin{itemize}
    \item[\faExclamationTriangle] Traditional security assumes a trustworthy OS — \alert{assumption broken}.
    \item[\faLock] Need isolation of critical code/data \textbf{even if the OS is compromised}.
\end{itemize}
\vfill
%\scriptsize\textit{This sets the stage for Trusted Execution Environments (TEEs).}
\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=1.2cm and 1cm, every node/.style={font=\small}]
    \node[draw, fill=orange!30, minimum width=2.5cm, minimum height=1cm, rounded corners] (app) {\textbf{App}};
    \node[draw, fill=gray!20, below=of app, minimum width=3cm, minimum height=1cm, rounded corners] (os) {\textbf{Untrusted OS}};
    \node[draw, fill=blue!15, right=3cm of app, minimum width=2.5cm, minimum height=1cm, rounded corners] (enclave) {\faLock\ \textbf{Enclave}};

    \draw[->, thick, red] (os) -- (app) node[midway, left, align=left, font=\footnotesize] {Possible interference\\ or data snooping};
    \draw[->, thick, keystoneblue] (app) -- (enclave) node[midway, above, font=\footnotesize] {Trusted Execution};
    \draw[->, thick, dashed, red] (os) -- (enclave) node[midway, below, font=\footnotesize] {Attempted attack blocked};
\end{tikzpicture}
\caption{Enclaves isolate computation from a compromised OS}
\end{figure}
\end{frame}

@@

\begin{frame}{Benchmarking Tools}
    \small
    \begin{itemize}
        \item \textbf{Dhrystone:} Measures CPU performance (DMIPS).
        \item \textbf{CoreMark:} Focuses on integer performance, used in embedded systems benchmarking.
        \begin{block}{Execution Modes}
        \begin{itemize}
            \item \textbf{Native Execution:} Benchmark runs as a normal process.
            \item \textbf{Enclave Execution:} Benchmark runs within a Keystone enclave.
        \end{itemize}
        \end{block}
    \end{itemize}
    %\vspace{0.5em}
    %\uncover<1->{\textbf{Key Focus:} These tools provide insights into both CPU-bound and system-level performance.}
\end{frame}

@@

% Legend
\begin{scope}[xshift=0cm,yshift=-1.5cm]
\draw[fill=keystoneblue!60] (0,0) rectangle (0.4,0.4) node[right=0.5cm] {Contiguous region};
\draw[fill=warmaccent!60] (5,0) rectangle (5.4,0.4) node[right=0.5cm] {Fragmented region};
\draw[pattern=north east lines,pattern color=black] (0,-0.6) rectangle (0.4,-0.2) node[right=0.5cm] {Wasted padding};
\draw[fill=gray!50,opacity=0.4] (5,-0.6) rectangle (5.4,-0.2) node[right=0.5cm] {PMP entry count bar};
\end{scope}

@@

% Thank You Slide with visual focus
\begin{frame}
\centering
\Huge\textbf{Thank You}\\
\large Questions? \faQuestionCircle
\vfill
%\small Universität zu Lübeck — Master’s Thesis Defense
\includegraphics[height=1cm]{uzl-logo.pdf}
\end{frame}
\end{document}
